---
title: "rmd"
author: "Minsik Kim"
date: "8/2/2021"
output: html_document
---
***Minsik Project***

Loading the data
``` {r}
data<-read.csv("activity.csv")
head(data)
```


Groupping data

```{r, echo=FALSE}
library(dplyr)
by_date<- data %>% group_by(date)
```

Total number of steps taken per day  
Making histogram of total number of steps taken per day  
Mediam and mean value of total number of steps taken per day  


```{r}
total_date<-by_date %>% summarise( steps = sum (steps))
total_date
barplot(total_date$steps~total_date$date)
hist(total_date$steps, breaks = 30)
summary(total_date$steps)
```


Calculating mean values of steps each day
noNA was introduced as NA influences summarise function
Plotting average value of steps, at specific time interval
Maximum value of steps can be found using subseting the max value

```{r}

by_date %>% summarise( steps = mean (steps))
noNA<-subset(data, is.na(data$steps)==F)
by_interval<-noNA %>% group_by(interval)
dummy<-by_interval %>% summarise (steps = mean (steps))
plot(dummy, type = "l")
subset(dummy, dummy$steps==(max(dummy$steps)))
```

Total numnber of missing values

```{r}
sum(is.na(data$steps))
sum(is.na(data$date))
sum(is.na(data$interval))
#avg value for the specific interval will be used to fill NAs.
avg_interval<-by_interval %>% summarise( steps = mean (steps))
avg_interval
#Making a fuction whose NAs were replaced with average values
replace<-function(data){
        for (i in 1:length(rownames(data))){
          if(is.na(data[i,1])){
              data[i,1]<-subset(avg_interval$steps, avg_interval$interval==(data[i,3]))
          }
          
        }
        return(data)
}
new_data<-replace(data)
head(new_data)
```

Grouping new data by date

```{r}

new_group<-group_by(new_data, date)
new_data2<-new_group %>% summarise(steps = sum(steps))
#Making histogram with modified data
hist(new_data2$steps, breaks = 30)
summary(new_data2)
```

Organizing data into weekend and weekday file. 

```{r}
new_data$day<-weekdays(as.Date(new_data$date))
new_data$week<-NA
for(i in 1:length(row.names(new_data))){
  if (new_data[i,4] == "Sunday"){
    new_data[i,5]<-"weekend"
  }
  else if (new_data[i,4] == "Saturday"){
    new_data[i,5]<-"weekend"
  }
  else {
    new_data[i,5]<-"weekdays"
  }
}
weekday_data<-subset(new_data, new_data$week=="weekdays")
weekend_data<-subset(new_data, new_data$week=="weekend")
#weekday data
weekday_by_interval<- weekday_data %>% group_by(interval)
weekday_profile<-weekday_by_interval %>% summarise( steps = mean (steps))
#weekend data
weekend_by_interval<- weekend_data %>% group_by(interval)
weekend_profile<-weekend_by_interval %>% summarise( steps = mean (steps))
```


Making plot

```{r}
par(mfrow = c(1,2))
plot(weekday_profile, type = "l")
title(main= "Pattern during weekdays")
plot(weekend_profile, type = "l")
title(main= "Pattern during weekends")
```
